### 2013-05-15 doc/pc-ijbc.tex
pc-ijbc: add logtime.gp, editing

\documentclass{ws-ijbc}
\usepackage{multirow}
\usepackage{centernot}
%\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{calc}
\usepackage{cancel} % strike out (cancel) in math mode
\usepackage{ws-rotating}
\begin{document}

\newcommand{\odd}{\mathrm{odd}}

% bold font for vectors
\newcommand{\vct}[1]{\mathbf{#1}}
% the cycle and sum operator
%\newcommand{\C}{\mathcal{C}}

% center column in tabularx
\newcolumntype{C}{>{\centering\arraybackslash}X}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}
\newcolumntype{R}{>{\raggedleft\arraybackslash}X}

\newcommand{\vx}{\vct x}
\newcommand{\vxn}[1]{\vx_{#1}}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\LB}{L_\Pset}
\newcommand{\lam}{(\lambda)}
\newcommand{\Rlam}{(R, \lambda)}
\newcommand{\rlam}{(r, \lambda)}

%\newcommand{\eq}{}
%\newcommand{\eqs}{}
\newcommand{\eq}{Eq.}
\newcommand{\eqs}{Eqs.}
\newcommand{\req}[1]{(\ref{eq:#1})}
\newcommand{\refeq}[1]{\eq\,\req{#1}}
\newcommand{\refeqs}[1]{\eqs\,\req{#1}}
\newcommand{\reqsub}[2]{(\ref{eq:#1}#2)}
\newcommand{\refeqsub}[2]{\eq\,\reqsub{#1}{#2}}
\newcommand{\refeqssub}[2]{\eqs\,\reqsub{#1}{#2}}

\newcommand{\refthm}[1]{Theorem \ref{thm:#1}}
\newcommand{\refthms}[1]{Theorems \ref{thm:#1}}
\newcommand{\refsec}[1]{Section \ref{sec:#1}}
\newcommand{\refsecs}[1]{Sections \ref{sec:#1}}
\newcommand{\refapd}[1]{Appendix \ref{apd:#1}}
\newcommand{\reftab}[1]{Table \ref{tab:#1}}
\newcommand{\reftabs}[1]{Tables \ref{tab:#1}}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\reffigs}[1]{Figs. \ref{fig:#1}}

% ws-ijbc doesn't like the following lines
%\newtheorem{theorem}{Theorem}
%\newenvironment{remark}[1][1]%
%{\par\noindent\textbf{Remark #1.} }{\medskip}

\catchline{}{}{}{}{}

\title{Cycles of the logistic map}
\author{Cheng Zhang}
%\affiliation{Applied Physics Program and Department of Bioengineering, \
%Rice University, Houston, TX 77005}
\address{Applied Physics Program \& Department of Bioengineering,
Rice University, Houston, TX 77005, USA}

\maketitle

\begin{history}
\received{}
\end{history}

\begin{abstract}
The onset and bifurcation points of the $n$-cycles of
  a polynomial map are located
  through a characteristic equation
  connecting the cyclic polynomials of the cycle points.
The minimal polynomials of the critical parameters
  of the logistic, H\'enon, and cubic maps are obtained
  for $n$ up to 13, 9, and 8,
  respectively.
\end{abstract}

\keywords{Logistic map; H\'enon map; cubic map; cycles; exact polynomials.}

%\maketitle

% switch text to two-column
\twocolumn

%
%
\section{Introduction}
%
%

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
The iterated sequence $x_1$,
  $x_2 = f(x_1)$,
  $x_3 = f(x_2)$, $\ldots$
can be conveniently visualized on the cobweb plot,
  as shown in \reffig{cobweb}.
%
Starting from $(x_1, x_1)$ on the diagonal,
  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
  where $y = f(x_k) = x_{k+1}$;
the next horizontal arrow then reflects $(x_k, y)$ to
  $(y, y) = (x_{k+1}, x_{k+1})$,
  the starting point of the next iteration.
%
%
\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:cobweb}
  Cobweb plot of the logistic map.}
\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a \emph{fixed point} (including infinity),
      e.g., \reffig{cobweb}(a);
  (ii) a periodic \emph{cycle},
      e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic trajectory,
      e.g., \reffig{cobweb}(f).
We will focus on the first two cases here.

A fixed point is a solution of $x^* = f(x^*)$.
%
It is stable if a small deviation from $x^*$
  gets reduced by successive iterations.
%
For a differentiable $f$,
  this is equivalent to $|f'(x^*)| \le 1$
  \cite{strogatz}.

An $n$-cycle is a sequence $x_1, \dots, x_n$,
  such that $n$ is the smallest positive integer
  that satisfies $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Obviously, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$
%
\big[but the converse is untrue, for a fixed point of $f^n$
  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz}.
%
%
%
The cycle and its stability depend on the parameter $r$.
%
The problem is to identify
  all regions of $r$ that allow stable $n$-cycles,
  which has been the objective of several previous
  studies
  \cite{brown1, brown2, stephenson1, stephenson2, stephenson3,
  saha, bechhoefer, gordon, burm, zhang, bailey1, bailey2, kk1, lewis}.
Below we present a simple algorithm to do so.

\section{\label{sec:logmap}Logistic map}

We will illustrate the algorithm on the logistic map \cite{may, strogatz}
  defined in \refeq{logmap}.
If $r$ is real,
  we will find windows $(r_a, r_b)$,
    within which stable $n$-cycles can exist.
%
%The window boundaries can be classified
%  to onset and bifurcation points according to \refeq{der};
Here, if $r > 0$, then
  $r_a$ ($r_b$) are the onset (bifurcation) points.
%
There are generally multiple such windows for a single $n$,
  but all onset points satisfy the same polynomial equation,
  and all bifurcation points another.
%
Our goal is thus to find the two polynomials for a given $n$.

To simplify the calculation, we first change variables \cite{mandelbrot, brown1, brown2, saha} as
%
%\begin{equation}
\[
    x^{\mathrm{new}} \equiv r(x^{\mathrm{old}} - 1/2),
    \quad R \equiv r(r-2)/4,
\]
%\label{eq:chvars}
%\end{equation}
%
and rewrite the map, in terms of $x^{\mathrm{new}}$, as
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv R - {x_k}^2.
  %\tag{$\ref{eq:logmap}'$}
\label{eq:logmaps}
\end{equation}
%
%\refeq{logmaps} simplifies \refeq{logmap} as $x_k$ now only occurs once.
%
We will solve the cycle boundaries as zeros of some polynomials of $R$,
and the corresponding polynomials for $r$ can be obtained by $R \rightarrow r(r-2)/4$.
%
%the numerical $r$ values can always be obtained from $r = 1 + \sqrt{1 + 4R}$.
%

\subsection{Overall plan}

Since the $n$-cycles points constitute a subset of the fixed points of $f^n$,
we will first find
  the polynomials at the stability boundaries
  for the fixed points of $f^n$
  (\refsecs{cyclic}-\ref{sec:examples}).
The contributions from shorter $d$-cycles ($d|n$)
  are lower-degree factors of the polynomials,
  and can be readily removed (\refsec{primfac}).

At the first glance, the polynomials of the fixed points of $f^n$
  can be obtained by brute force.
We can solve
  \refeqs{logmaps} and express $x_1, \ldots, x_n$
  in terms of $R$,
  and then plug the solution into \req{der}.
The result would contain $R$ only,
  and is therefore the answer.
But since \refeqs{logmaps} are nonlinear,
  it is nontrivial to reduce the final equation of $R$
  into a polynomial one.
Nonetheless, on a computer, one can
  construct a polynomial Gr\"obner basis \cite{kk1},
  which gradually eliminates $x_k$ out of the $n+1$ equations.
As a special case, if $x_2, \dots, x_n$ are eliminated first,
  the final elimination of $x_1$ can be accomplished
  by setting the resultant of $f^n(x_1) = x_1$ and
  \refeq{der} or $\frac{d}{dx}f^n (x_1) = \pm 1$
  to zero \cite{burm},
  which furnishes the needed polynomial equation of $R$.
%
These approaches, however, do not
  exploit the cyclic structure of \refeqs{logmaps},
  and can thus be improved by the following alternative.

Instead of solving \refeqs{logmaps} for $x_k$,
  we will derive a set of \emph{homogeneous linear} equations
  $\sum_j A_{ij} C_j = 0$
  regarding cyclic polynomials $\{C_j\}$ of $x_k$
  (an example of a cyclic polynomial is $x_1 x_2 + x_2 x_3 + \dots + x_n x_1$).
%
Now the matrix formed by the coefficients $A_{ij}$ of the equations
  must have a zero determinant,
  for the cyclic polynomials $\{C_j\}$ are not zeros altogether.
%
Thus, the zero-determinant condition $|\mathbf A| = 0$ gives
  the needed polynomial equation of $R$,
whose roots contain all fixed points of $f^n$.
%

\subsection{\label{sec:cyclic}Cyclic polynomials}

A cyclic polynomial is an invariant
  under the cycling of variables
  $x_1 \rightarrow x_2, x_2 \rightarrow x_3,
  \ldots, x_n \rightarrow x_1$,
  e.g., $a = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$
  and $b = x_1 x_3 + x_2 x_4$ (for $n = 4$).
%
It is, however, unnecessarily symmetric
  (i.e., invariant under the exchange of any two $x_k$ and $x_{j}$),
  e.g., $a$ and $b$ are not symmetric,
  although $a + b$ is.

A cyclic polynomial can be generated by summing over
  distinct cyclic versions of a simpler polynomial of $x_k$, or a \emph{generator}.
For example,
$x_1 x_2$ is a generator of $a$;
$x_1 x_3$ is that of $b$;
and
$x_1 x_2 + \frac{1}{2} x_1 x_3$ is that of $a + b$.
Note that the coefficient before $x_1 x_3$
  is 1 in the second case
  for there are only two distinct versions,
  but $\frac{1}{2}$ in the third case
  for there are four. % distinct versions of the generator.

Cyclic polynomials
  generated from monomials of unit coefficient
  (such as $a$ and $b$, but not $a + b$ in the above example)
  can be labeled as follows.
We write the monomial generator as
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  and form a sequence $p$ of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s.
The sequence $p$ serves as the label of the cyclic polynomial $C_p(\vxn{n})$,
  where $\vxn{n} \equiv \{x_1, \ldots, x_n\}$
[we will, however, drop the subscript ``$_n$'' or
  even the argument ``$(\vxn{n})$'' entirely when it is clear].
For example,
  $C_{12}(\vx)  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$ and
  $C_{112}(\vx) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
In case the cyclic polynomial has several generators,
  we pick the one with the smallest $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ instead of $x_2 x_3$ for $C_{12}(\vx)$
  because $12 < 23$ ($n \ge 3$).
Finally, we add $C_0(\vx) \equiv 1$ for completeness.
%
%
%

\subsection{Square-free reduction}

A cyclic polynomial is \emph{square-free}
  if it has no square or higher powers of $x_k$,
e.g., $C_{12}(\vx) = x_1 x_2 + \cdots$ is square-free,
  $C_{112}(\vx) = {x_1}^2 x_2 + \cdots$ is not.
%see \reftab{sqrfreepoly} for more examples.
The label $p$ of a square-free polynomial
  has no repeated index.
We denote the set of all square-free $p$ by
$\Pset = \{0, 1, 12, 13, \ldots, 123, 124, \ldots, 12\dots n\}$,
and its size by $\NB$.
% is the number of square-free cyclic polynomials.
%
%
The set of square-free cyclic polynomials
  furnishes a basis for expanding cyclic polynomials:

\begin{theorem}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx)$
  formed by the $n$-cycle points
  $\vx = \{x_1, \ldots, x_n\}$
  is a linear combination of
  the square-free cyclic polynomials $C_p(\vx)$:
\[
  K(\vx) = \sum_{p \in \Pset} f_p(R) \, C_p(\vx),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of indices of all square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{theorem}
%
%
\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx)$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since both the replacement rules and the original polynomial are cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vx) \equiv 1$.
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only new variable introduced by the substitutions.
\end{proof}
%
%
For example, the cyclic polynomial
  $K(\vxn{2}) = {x_1}^2 \, x_2 + {x_2}^2 \, x_1$
can be written as $K = (R + 1) \, C_{1} - 2 R \, C_0$
for
${x_1}^2 \, x_2 = R \, x_2 - {x_2}^2 = R \, x_2 - R + x_1$ and
${x_2}^2 \, x_1 = R \, x_1 - R + x_2$.

\refthm{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the square-free ones.
Below we show that at the onset and bifurcation points,
  the square-free cyclic polynomials $C_p$ are themselves
  linearly connected by $\NB$ homogeneous equations.
The matrix of the connecting coefficients thus has a vanishing determinant,
  which furnishes the desired polynomial equation.

\subsection{\label{sec:algo}Fixed points of $f^n$}

We first observe that the derivative of $f^n$
is a cyclic polynomial:
%
%
\begin{equation}
  \Lambda(\vx)
   = f'(x_1) \dots f'(x_n)
   = (-2)^n x_1 \dots x_n.
\label{eq:logder}
\end{equation}
%
%
Since $C_p(\vx)$ is also cyclic,
  so is $\Lambda(\vx) \, C_p(\vx)$,
The latter can be expanded by \refthm{sqrfree} as
\begin{equation}
  \Lambda(\vx) \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx),
\label{eq:xcp0}
\end{equation}
where $T_{pq}(R)$ is a polynomial of $R$,
and $p, q \in \Pset$.

By \refeq{der},
  $\Lambda(\vx)$ equals to $\lambda = +1$ or $-1$
  at the onset or bifurcation point,
  respectively;
  so \refeq{xcp0}
  becomes a homogeneous linear equation of
  $C_p(\vx)$:
\begin{equation}
  \lambda \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx),
\label{eq:xcp}
\end{equation}
%
%
%where $\lambda$ is $+1$ and $-1$ at the onset and bifurcation points, respectively;
or in matrix form,
\begin{equation}
  \big[ \lambda \, \vct I - \vct T(R) \big] \, \vct C = 0,
\tag{\ref{eq:xcp}$'$}
\end{equation}
%
%
where
$\vct I$ is the $\NB \times \NB$ identity matrix,
$\vct T(R) = \{T_{pq}(R)\}$ is an $\NB \times \NB$ matrix,
and
$\vct C = \{C_p(\vx)\}$ is an $\NB$-dimensional column vector.

Since a set of homogeneous linear equations
  has a non-trivial solution
  (i.e., $C_p(\vx)$ do not vanish altogether)
  only if the determinant of the coefficient matrix
  is zero, we have
\begin{equation}
  A_n\Rlam \equiv \big| \lambda \, \vct I - \vct T(R) \big| = 0.
\label{eq:xr}
\end{equation}
%
Now $A_n\Rlam$, after we expand the determinant,
  is a polynomial of $R$ and $\lambda$,
  the equation is the desired answer to our problem.
%
We attach the subscript $n$ to $A\Rlam$ because
  it will later be used with $A_d\Rlam$ of divisors $d$ of $n$.
%
To summarize, we have
\begin{theorem}
  At the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx)$
  of the cycle points $\vx = \{x_k\}$
  are linearly related by % \refeq{xcp},
  $\lambda \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx)$,
  with $\lambda$ being $+1$ and $-1$, respectively,
  and $T_{pq}(R)$ being the coefficients
  from the square-free reduction of the cyclic polynomial
  $C_p(\vx) \, \prod_k f'(x_k)$. %[\refeq{logder}].
  Thus, the fixed point of $f^n$
  is marginally stable when $R$ satisfies
  characteristic equation $\big| \lambda \, \vct I - \vct T(R) \big| = 0$.
  \label{thm:main}
\end{theorem}

%\begin{remark}
More generally, the algorithm can trace a contour of marginal stability of $R$
  in the complex plane
  for complex cycles $\vx$
  if $\lambda$ is set to $\exp(i\theta)$ with $\theta \in [0, 2\pi]$.
  The region enclosed by the contour corresponds to
    a bulb in the inverted Mandelbrot set
  (it follows from the transformation $z_k = -x_k, c = -R$, which
  reduces \refeq{logmaps} to the quadratic map $z_{k+1} = c + {z_k}^2$
  used in the Mandelbrot set \cite{mandelbrot}),
  see ref. \cite{stephenson2, stephenson3}. % and \reffig{logbulb}.
%\end{remark}

%\begin{remark}
%  With $\lambda = 0$,
%  the algorithm yields the polynomial of the superstable point,
%  at which the initial deviation from a cycle point vanishes
%  to the linear order after $n$ iterations of $f$.
%  But we have a better algorithm in this case:
%  since at least one of the $x_k$ is zero by \refeq{logder},
%  then $f^n(x_k = 0) = x_{n+k} = 0$, which provides
%  the desired equation of $R$ \cite{strogatz}.
%\end{remark}

\subsection{\label{sec:examples}Examples}

We illustrate the above algorithm on several small-$n$ cases.
%
It is still helpful to have a mathematical software
  at hand to verify certain steps
  (e.g., in computing the determinants and their factorization).

For $n = 1$, we have
  $C_0 = 1$, $\Lambda = -2 \, C_1 = -2 \, x_1$, and
\[
  \Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right)
  =
  \left( \begin{array}{cc}
  0     & -2 \\
  -2R   &  2
  \end{array}\right)
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right).
\]
With $\Lambda \rightarrow \lambda$, the characteristic equation (\ref{eq:xr}) reads
\begin{equation}
0
  %=\left| \begin{array}{cc}
  %\lambda     & 2          \\
  %2 R         & \lambda - 2
  %\end{array} \right|
  = -4R - 2\lambda + \lambda^2.
  \tag{\ref{eq:xr}-1}
\label{eq:xr1}
\end{equation}
Thus, $R = (\lambda^2 -2\lambda)/4$;
  and the fixed point begins at $R_a = -1/4$ ($\lambda = +1$), and
  becomes unstable at $R_b = 3/4$ ($\lambda = -1$).

For $n = 2$, the cyclic polynomials
  are
$C_0 = 1$,
$C_1 = x_1 + x_2$,
$C_{12} = x_1 x_2$;
and $\Lambda = 4 \, x_1 x_2$.
Thus,
$\Lambda \, C_0 = 4 \, C_{12}$,
$\Lambda \, C_1 %= 4 \, {x_1}^2 \, x_2 + 4 \, x_1 \, {x_2}^2
  = -8 \, R \, C_0 + 4 \, (R + 1) \, C_1$
by the example after \refthm{sqrfree},
$\Lambda \, C_{12}
  = 4 \, (R - x_2) (R - x_1)$,
and
\[
\Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12}
  \end{array} \right)
 =
  \left( \begin{array}{ccc}
  0           & 0         & 4 \\
  -8R         & 4(R+1)    & 0 \\
  4R^2        & -4R       & 4
  \end{array} \right)
  \left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12}
  \end{array} \right).
\]
%
%
%
The characteristic equation \refeq{xr} reads
%
%
%
\begin{align}
0 &= (4R - 4 + \lambda) [(4 R-\lambda)^2 - 4 \lambda].
  \tag{\ref{eq:xr}-2}
\label{eq:xr2}
\end{align}
%
%
%
The first factor is responsible for the 2-cycle,
  while the second factor is for the unstable fixed points. % (see later).
Setting the former to zero yields $R = 1 - \lambda/4$;
so the onset ($\lambda = +1$) value is $R_a = 3/4$ ($r_a = 3$)
and the bifurcation ($\lambda = -1$) value is $R_b = 5/4$ ($r_b = 1+\sqrt 6$).
%
%Note that the onset point of the 2-cycle
%  is located at $R = 3/4$,
%  where the fixed point bifurcates \cite{strogatz}.
%  \big[compare \reffigs{cobweb}(a) and (b)\big].

%The $n = 3$ case is nontrivial and has sparked several different
%  elementary derivations
%  \cite{brown1, saha, bechhoefer, gordon, burm, zhang}.
%For the current algorithm,
%%the cyclic polynomials are
%%$C_0 = 1$,
%%$C_1 = x_1 + x_2 + x_3$,
%%$C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$,
%%$C_{123} = x_1 x_2 x_3$,
%%and
%$\Lambda = -8 \, C_{123}$,
%%
%and the square-free reduction yields
%%
%\begin{small}
%\begin{align*}
%\Lambda
%\left( \begin{array}{c}
%  C_0 \\
%  C_1 \\
%  C_{12} \\
%  C_{123}
%\end{array} \right)
%=
%\left( \begin{array}{cccc}
% 0            & 0                 & 0               & -8 \\
% -24 R        & 8(R+1)            & -8R             & 0 \\
% 24 R^2       & -8R(R+2)          & 8(R+1)          & 0 \\
% -8R^3        & 8R^2              & -8R             & 8
%  \end{array} \right)
%\left( \begin{array}{c}
%  C_0 \\
%  C_1 \\
%  C_{12} \\
%  C_{123}
%\end{array} \right),
%\end{align*}
%\end{small}
%%
%%
%%
%and \refeq{xr} reads:
%\begin{align}
%  0
%&= -\big[
%  64R^3 - 128 R^2 - 8 (\lambda - 8) R - (\lambda -8)^2
%  \big] \notag \\
%&\qquad
%  \big(
%  \lambda^2 - 8\lambda - 24 R\lambda - 64 R^3
%  \big).
%  \tag{\ref{eq:xr}-3}
%\label{eq:xr3}
%\end{align}
%%
%Using the first factor, we find at the onset point
%  $\lambda = 1$,
%  $\left(
%  R-\frac74
%  \right)
%  \left(
%  R^2-\frac14 R + \frac{7}{16}
%  \right)=0$
%  and its only real solution is $R_a = 7/4$ ($r_a = 1+\sqrt 8$).
%%
%At the bifurcation point $\lambda = -1$,
%  the equation $R^3-2R^2 + \frac{9}{8} R -\frac{81}{64}=0$
%  yields
%  $R_b = \frac14
%      \left(
%      \frac{8}{3} +
%        \sqrt[3]{\frac{1915}{54} - \frac52\sqrt{201}}
%       +\sqrt[3]{\frac{1915}{54} + \frac52\sqrt{201}}
%      \right)$,
%whose corresponding $r = 1+\sqrt{1+4R}$
%is identical to that in ref. \cite{gordon, burm}.
%%

For the $n = 4$ case
\cite{brown1, stephenson1, stephenson3, bailey1, bailey2, kk1, lewis},
%$C_0 = 1$,
%$C_1 = x_1 + \cdots + x_4$,
%$C_{12} = x_1 x_2 + \cdots + x_4 x_1$,
%$C_{13} = x_1 x_3 + x_2 x_4$,
%$C_{123} = x_1 x_2 x_3 + \cdots + x_4 x_1 x_2$,
%$C_{1234} = x_1 x_2 x_3 x_4$,
%and
%$\Lambda = 16 \, C_{1234}$.
%
we only list the final characteristic equation \eqref{eq:xr}
after the square reduction:%
%(with six square-free cyclic polynomials):
%
\begin{small}
\begin{align}
0
%& = \left|
%\begin{array}{cccccc}
% \lambda  & 0               & 0                 & 0     & 0     & -16 \\
% 64 R     & \lambda-16(R+1) & 16 R              & 0     & -16R  & 0 \\
% -64 R^2  & 16R(R+2)        & \lambda-16(R^2+1) & -32 R & 16R   & 0 \\
% -32 R^2  & 16R(R+1)        & -16R              & \lambda-16(R^2+1)  & 0 & 0 \\
% 64 R^3   & -16R^2(R+3)     & 16R(R+2)        & 32 R (R+1)       & \lambda-16(R+1) & 0 \\
% -16R^4   & 16R^3           & - 16R^2           & -16R^2            & 16R & \lambda - 16
%\end{array}
%\right | \nonumber  \\
& =
    \big[
    4096 R^6-12288 R^5 + 256 (\lambda + 48) (R^4- R^3) \notag \\
& \quad
    -16 (\lambda + 32)(\lambda - 16) R^2
      -(\lambda - 16)^3 \big] \notag \\
& \quad
    \big[ 16 (R-1)^2 - \lambda \big]
 \, \big[(16 R^2 + \lambda)^2 - 16 (2R+1)^2 \lambda \big],
  \tag{\ref{eq:xr}-4}
\label{eq:xr4}
\end{align}
\end{small}
%
%
%
The first (sextic) factor, responsible for the 4-cycle,
can be factorized as
$ (
    4R - 5
  )
\, \big[
  (4R + 1)^2 + 4
  \big]
\, \big[
  (4R - 3)^3 - 108
  \big] =0$
at the onset point $\lambda = 1$.
%
%
%
It has two real roots:
$R_a' = 5/4$
\big($r_a' = 1+\sqrt{6} \approx 3.4495$\big)
for the cycle from period-doubling the 2-cycle
  \big[compare \reffigs{cobweb}(b) and (d)\big],
and
$R_a = (3+\sqrt[3]{108})/4$
\big($r_a = 1+\sqrt{4+\sqrt[3]{108}} \approx 3.9601$\big)
for an original cycle \big[\reffig{cobweb}(e)\big].
%
At the bifurcation point ($\lambda = -1$), we have
$4096 R^6 - 12288 R^5 + 12032 (R^4 - R^3)
  + 8432 R^2 + 4913 = 0$
%which upon $R \rightarrow r(r-2)/4$ yields the same polynomial
%obtained previously
\cite{stephenson1, stephenson3, bailey2, kk1, lewis}.
%
%The only two positive roots $r_b \approx 3.9608$
%and $r_b' \approx 3.5441$ of the sextic equation
%correspond to $r_a$ and $r_a'$ respectively.
%
%Since the 4-cycle problem were tackled mainly on computers
%  in the previous studies,
%  we also supply a compact elementary derivation
%  in the Appendix.

\subsection{\label{sec:primfac}$n$-cycle polynomial}

The characteristic polynomial $A_n\Rlam$ in \refeq{xr}
  is derived for the fixed points of $f^n$.
Thus, it represents not only the $n$-cycles,
  but also cycles of shorter periods $d$ ($d|n$).
%
Fortunately, the contributions from different cycles
  are separate factors of $A_n\Rlam$.
For a large $n$,
  the factor $P_n\Rlam$ that represents all $n$-cycles
  can be easily recognized
  as the largest factor of $A_n\Rlam$.
%
More systematically, we have

%\begin{theorem}
%  The polynomial $P_n\Rlam$ of all $n$-cycles
%  is a factor of the characteristic polynomial $A_n\Rlam$,
%  and can be computed as
  \begin{equation}
    P_n\Rlam
    = \prod_{cd = n} B_d^c\Rlam^{\mu(c)},
  \label{eq:primfac}
  \end{equation}
where
  $B_d^c\Rlam \equiv \prod_{k=1}^c A_{d}(R, e^{2k\pi i/c} \lambda^{1/c})$,
  $\lambda^{1/c}$ is a complex $c$th root of $\lambda$.
%\end{theorem}
%
%
%\begin{remark}
%Here, ``minimal'' means that the polynomial contains all zeros.
%It happens to be irreducible for a general $\lambda$.
%At $\lambda = 1$, the polynomial of $R$, however, can be factorized (see later).
%\end{remark}
%
%\begin{remark}
The M\"obius function $\mu(n)$ is $(-1)^k$
  if $n$ is the product of $k$ distinct primes,
  or 0 if $n$ is divisible by a square of a prime.
%Thus,
The first few $\mu(n) = 1, -1, -1, 0, -1, 1, -1, \ldots,$
  from $n = 1$
%The $\mu(n)$ is useful for inversion:
%$g(n) = \sum_{d|n} \mu(n/d) \, h(d)$
%if and only if $h(n) = \sum_{d|n} g(d)$
 \cite{hardy}.
%\end{remark}
%

Since the discussions in this and the next sections
  concern $\lambda$ instead of $R$,
  we will drop the argument $R$ in $P_n\Rlam$ and $A_n\Rlam$, etc.
%\begin{remark}
%$B_{d,c}\lambda$ is a polynomial of $\lambda$.
Despite the argument $\lambda^{1/c}$,
  the product $\prod_{k=1}^c A_d(e^{2k\pi i/c} \lambda^{1/c})$,
  hence $B_d^c\lam$,
  is free from radicals of powers of $\lambda^{1/c}$,
  for it is invariant under $\lambda \rightarrow e^{2\pi i} \lambda$.
Thus, $B_d^c\lam$ is a polynomial of $\lambda$,
and $\deg_\lambda B_d^c\lam = \deg_\lambda A_d\lam$.
Particularly, $B_n^1\lam = A_n\lam$.
%
%\end{remark}

%
Let us apply \refeq{primfac}
to the examples in \refsec{examples}.
%%
For $n = 1$,
%there is no irrelevant factor in \refeq{xr1} and
$P_1(\lambda) = B_1^1(\lambda) = A_1(\lambda)
  = \lambda^2 - 2\,\lambda - 4\,R$.
%
%
%
For $n = 2$, since
$B_1^2(\lambda)
=
A_1(\sqrt{\lambda})
A_1(-\sqrt{\lambda})
=
(4R-\lambda)^2 -4\lambda$,
%
$P_2(\lambda) = A_2(\lambda) \, B_1^2(\lambda)^{-1} = 4R - 4 + \lambda$.
%
%
For $n = 4$,
%we have
%$16 (R-1)^2 - \lambda
%= (4R - 4 + \sqrt{\lambda})(4R - 4 - \sqrt{\lambda})$
%and
%$
%(16 R^2 + \lambda)^2 - 16 (2R+1)^2 \lambda
%=
%\big[(4R - \sqrt{\lambda})^2  - 4\sqrt{\lambda}\,\big]
%\big[(4R + \sqrt{\lambda})^2  + 4\sqrt{\lambda}\,\big]$.
%Thus,
we can show that
the last two factors of \refeq{xr4}
can be written as
$B_2^2(\lambda) = A_2(\sqrt{\lambda}) A_2(-\sqrt{\lambda})$,
and
%
$P_4(\lambda)
    = A_4(\lambda) \, B_2^2(\lambda)^{-1}$
indeed the yields the sextic factor.
%$    = 4096 R^6-12288 R^5+ 256 (\lambda + 48) (R^4- R^3)
%    -16(\lambda + 32)(\lambda - 16) R^2 - (\lambda-16)^3$.
Note, $B_1^4(\lambda)$ is unused for $\mu(4) = 0$.

%It can be, however, problematic,
%   if $P_n(R, \lambda)$ is solved for $\lambda = 1$
%   instead of a general $\lambda$,
%   for $P_n(R, 1)$ itself can be further factorized.
%
Due to the technical nature of the derivation and subsequent discussions,
  the reader may wish to skip the rest of \refsec{logmap}
  on first reading.

\subsection{\label{sec:degprimfac}Counting cycles}

%We will prove \refthm{primfac} by the following steps.
%
We now prove \refeq{primfac}.
%
To do so, we first find the degrees in $\lambda$ of
  $A_n\lam$ [\refeq{necklace}] and $P_n\lam$ [\refeq{lyndon}].
By comparing the degrees,
  we then show that each $P_d\lam$ with $d|n$,
  after some transformation,
  contributes one polynomial factor to $A_n\lam$
  [\refeq{prod}],
  and the inversion of the relation
  yields \refeq{primfac}.

%\subsubsection{Number of the square-free cyclic polynomials}

First, to count the square-free cyclic polynomials,
  we notice a one-to-one mapping
  from the square-free cyclic polynomials
  to the binary necklaces.
%
A binary necklace
  is a nonequivalent binary $0$-$1$ string.
%
Two strings are equivalent if they differ only by a circular shift.
%
For example, for $n = 3$, there are $2^3 = 8$ binary strings,
but only four necklaces: 000, 001, 011 and 111,
since 010 and 100 are equivalent to 001,
so are 110 and 101 to 011.
%
The mapping is constructed as follows.
%
Given a cyclic polynomial, if the generator contains $x_k$,
  the $k$th character from the right of the necklace is 1,
  otherwise 0,
e.g. the above four necklaces correspond to
  $C_0$, $C_1$, $C_{12}$ and $C_{123}$, respectively.
%For example, if $n = 3$,
% the square-free polynomials for $000$, $100$, $110$ and $111$ are
%  $C_0 = 1$ (generator: 1),
%  $C_1 = x_1 + x_2 + x_3$ (generator: $x_1$),
%  $C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$ (generator: $x_1 x_2$)
%  and $C_{123} = x_1 x_2 x_3$ (generator: $x_1 x_2 x_3$), respectively.
Thus,
%
%\begin{theorem}
the number $\NB(n)$ of the square-free cyclic polynomials $C_p$,
  %$(\vx)$ formed by the cycle points $\vx = \{x_1, \ldots, x_n\}$,
  is given by
\begin{equation}
  %\deg_\lambda A_n\lam =
  \NB(n) = (1/n) \textstyle\sum_{d|n} \varphi(n/d) 2^{\,d},
\label{eq:necklace}
\end{equation}
%
%\label{thm:necklace}
%\end{theorem}
%
where $\varphi(m)$ is Euler's totient function,
which gives the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(3) = 2$ for 1 and 2,
  $\varphi(6) = 2$ for 1 and 5.
By \refeq{xr},
  $\NB$ is also $\deg_\lambda A_n$.
The first few $\NB(n)$ are
2, 3, 4, 6, 8, 14, 20, 36, 60, 108, \dots, % 188, 352, 632, \ldots,
  from $n = 1$.

%\subsubsection{Number of the $n$-cycles}

%\begin{theorem}
Next the polynomial $P_n\lam$ can be written as
\begin{equation}
  P_n\lam = \textstyle\prod_l [\lambda - \Lambda(\vxn{n}^{(l)})],
\end{equation}
where $\vxn{n}^{(l)}$ denotes the $l$th $n$-cycle.
%
Thus, $\deg_\lambda P_n\lam$
  is equal to the number $\LB$ of the $n$-cycles,
  which is given by
  \cite{hao, hao2, lutzky}.
\begin{equation}
  %\deg_\lambda P_n\lam =
  \LB(n) = (1/n)\textstyle\sum_{d | n} \mu(n/d) 2^{\,d}.
\label{eq:lyndon}
\end{equation}
%\label{thm:lyndon}
%\end{theorem}
%
%
%\end{proof}
%
The first few $\LB$ are
2, 1, 2, 3, 6, 9, 18, 30, 56, 99, \dots %186, 335, 630, \ldots,
from $n = 1$.

Incidentally, $\LB$ is also the number of aperiodic necklaces of length $n$
(e.g., for $n = 4$,
  ``$0011$'' is aperiodic, but ``$0101$'' is not due to the shorter period 2).
%
Since an $n$-necklace is either aperiodic itself or
  a repeated aperiodic $d$-necklace (with $d|n$),
  we have
\begin{equation}
  \NB(n) = \textstyle\sum_{d|n} \LB(d).
\label{eq:necklacelyndon}
\end{equation}
%
Since the left side is the degree in $\lambda$ of $A_n\lam$,
  which accommodates all fixed points of $f^n$,
  %\refeq{necklacelyndon} shows that
  each distinct $d$-cycle (with $d|n$)
  contributes one and only one zero to $A_n\lam$.
%
%
We may thus write
%\begin{theorem}
  \begin{equation}
    A_n\lam = \textstyle\prod_{d|n} Q_d^{n/d}\lam,
    \label{eq:prod}
  \end{equation}
where $Q_d^{n/d}(\lambda) \equiv \prod_l
  \big[\lambda - \Lambda\big(\vxn{n}^{(l)}\big) \big]$
  with $\vxn{n}^{(l)}$ being the $n$ points
  corresponding to the $l$th $d$-cycle.
%
In order words, $\vxn{n}^{(l)}$ is $\vxn{d}^{(l)}$
  repeated for $c = n/d$ times:
\begin{equation}
  \vxn{n}^{(l)} =
  \{
    x_1^{(l)}, \ldots, x_d^{(l)},
    \; \ldots, \;
    x_1^{(l)}, \ldots, x_d^{(l)}
  \}
  = \vxn{d}^{(l)} * c.
\notag
\end{equation}
%
It follows that
$\Lambda(\vxn{n})$ = $[\Lambda(\vxn{d})]^c$.
By factorizing $\lambda - \Lambda^c$
  in terms of complex roots of unity, we get
\begin{align}
Q_d^{c}\lam
%  &= \prod_l
%    \Big\{
%      \lambda - \big[\Lambda\big(\vxn{d}^{(l)}\big) \big]^{c}
%    \Big\}
%    \notag \\
  &=
    \prod_l
    \Big\{
      \prod_{k=1}^c \,
      \big[
          e^{2k\pi i/c} \lambda^{1/c}
          - \Lambda\big(\vxn{d}^{(l)}\big)
      \big]
    \Big\},
    \notag
\end{align}
%
Now changing the order of the two products yields
\begin{equation}
Q_d^{c}\lam
  =  \prod_{k=1}^c P_d\big(e^{2k\pi i/c} \lambda^{1/c}\big),
\label{eq:QP}
\end{equation}
%
%
Despite the presence of $\lambda^{1/c}$,
  $Q_d^c\lam$ is a polynomial of $\lambda$,
  thanks to the invariance
  $\lambda \rightarrow e^{2\pi i} \lambda$;
and
  $\deg_\lambda Q_d^c\lam
    = \deg_\lambda P_d\lam = \LB(d)$.

%Therefore, the product $\prod_{cd=n} Q_d^c\lam$
%  can differ from $A_n\lam$ only by
%  a multiple.
%Since $Q_n^1\lam = P_n\lam$,
%  and the coefficient of highest power of $\lambda$ is always
%  unity in $A_n\lam$ \big[see the definition \refeq{xr}\big],
%  we know by induction that the coefficients of the highest power of $\lambda$
%  in all $P_n\lam$ and $Q_d^c\lam$
%  are also unities.
%So the multiple is one, hence \refeq{prod}.

\refeq{primfac} can now be derived from
  \refeqs{prod} and (\ref{eq:QP}),
%\begin{small}
\begin{align*}
 B_n^m (\lambda)
% & = \prod_{r = 1}^{m} A_n(e^{2l\pi i/m} \lambda^{1/m}) \\
 & = \prod_{k' = 1}^m \prod_{c | n} \prod_{k=1}^c
      P_{n/c}\big(
        e^{2k\pi i/c + 2 k' \pi i/(m c)}
        \lambda^{1/(m c)}
      \big) \\
 & = \prod_{c | n} \prod_{k''=1}^{m c}
    P_{n/c}(e^{2k''\pi i/(m c)} \lambda^{1/(m c)}) \\
 &
  = \prod_{c | n} Q_{n/c}^{m c}(\lambda).
\end{align*}
%\end{small}
The M\"obious inversion yields
\begin{equation}
Q_n^m\lam
= \textstyle\prod_{d|n} \big[
    B_d^{m n/d}\lam
  \big]^{\mu(n/d)}.
\notag
%\tag{$\ref{eq:primfac}'$}
\end{equation}
\refeq{primfac} is the special case of $m = 1$,
for $Q_n^1 = P_n$.

\subsection{\label{sec:origfac}%Intersection of cycles and
  Factors at the onset point}

The $n$-cycle polynomial
  $P_n(R, \lambda)$
  can be further factorized at $\lambda = 1$.
%
This is caused by the intersection of an $n$-cycle
  and a shorter $d$-cycle ($d|n$, $d < n$),
  which forces the two to share orbit.
%
One can show that (1) such an intersection happens only
  at the onset of the longer $n$-cycle, where $\lambda = 1$,
%
  and (2) $d$ and $n$ are coprime \cite{blackhurst}.
%
The simplest example is the first bifurcation point
  at $R = 3/4$ for $d = 1, n = 2$,
  where the fixed point \refeq{xr1}
  undergoes period-doubling and becomes the 2-cycle \refeq{xr2}.
As a result, the $n$-cycle polynomial $P_n(R, \lambda)$
  has to accommodate the $d$-cycle one $P_d(R, \lambda')$ as a factor,
  with $\lambda'$ being a primitive $(n/d)\,$th root of $\lambda = 1$.

We can therefore distinguish \emph{original} cycles
  from those born out of the above branching.
Similar to \refeq{QP},
  the contributions from the non-original cycles
  born out of the $d$-cycles
  can be written as a cyclotomic product
  $\prod_{(k, c) = 1}
      P_{d}
        \left(
          R, e^{2k\pi i/c}
        \right)$,
where $k$ runs through integers from 1 to $c=n/d$,
  that are coprime to $c$.
With these contributions removed,
  we obtain the polynomial $S_n(R)$ responsible for the original cycles:
  %or the \emph{original factor} below,
  \begin{equation}
    S_n(R)
    = \frac
    {
      P_n(R, 1)
    }
    {
      \prod_{c d =  n, \; c > 1}
      %\left[ \,
        \prod_{(k, c) = 1}
      P_{d}
        \left(
          R, e^{2k\pi i/c}
        \right)
      %\right]
    },
  \label{eq:origfac}
  \end{equation}
%
%
%
Take the $n = 4$ case as an example, we have
$\prod_{k=1,3} P_1(R, e^{k\pi i/2})
%=(-4R-2i-1)(-4R+2i-1)
=16R^2+8R+5$
(for $c = 4$),
$P_2(R, -1) = 4R - 5$
(for $c = 2$),
and
$P_4(R, 1)
= (4R - 5) (16 R^2 + 8 R + 5)
  \bigl[
    (4R - 3)^3 - 108
  \bigr]$.
Dividing $P_4(R, 1)$ by the two factors yields
$S_4(R) = (4R-3)^3 - 108$.
%  whose only real root $R_a=(3+\sqrt[3]{108})/4$ corresponds to
%  the onset of the original cycle.
Note $R_a' = 5/4$ is not original,
  for it is born out of period-doubling the 2-cycle.
The two cycles are compared in \reffigs{cobweb}(d) and (e).

%We now prove \refthm{origfac}.
%Suppose $n = c \,d$, we have, from \refeq{logmaps},
%\[
%  x_{l+1} - x_{d+l+1} = - (x_l + x_{d+l}) (x_l - x_{d+l}).
%\]
%We apply the equation to $l = 1, \ldots, m$, and the product is
%\[
%  x_{m+1} - x_{m + d+1} =
%  (-1)^{m} \left[ \, \prod_{l=1}^{m} (x_l + x_{d+l}) \right]
%    (x_1 - x_{d+1}).
%\]
%We now set $m$ to $0, d, \ldots, (c-1)\,d$ in this equation,
%  add them together, eliminate $x_1 - x_{d+1}$
%  (which is nonzero in a cycle), and
%\begin{equation}
%  \sum_{c' = 0}^{c-1}
%    (-1)^{c' d} \prod_{l=1}^{c' d} (x_l + x_{d+l})
%   = 0.
%\label{eq:stair}
%\end{equation}
%%
%Note \refeq{stair} holds for every divisor $c$ of $n$ ($c > 1$). We now have
%
%
%
%\begin{theorem}
%An $n$-cycle and a shorter $d$-cycle ($d|n$, $d< n$)
%  intersect only at the onset of the $n$-cycle,
%%
%and $\prod_{k=1}^{d} f'(x_k) = (-2)^d x_1 \dots x_d$ is
%a \emph{primitive} $(n/d)$th root of unity there.
%\label{thm:cbifur}
%\end{theorem}
%
%
%\begin{proof}
%At the intersection of the $n$- and $d$-cycles,
%  $x_l$ repeats itself after $d$ steps, so
%  $x_{d + l} = x_l$;
%and \refeq{stair} becomes,
%%
%\begin{equation}
%  1 + q + \dots + q^{c-1} = 0,
%\label{eq:cbifur}
%\end{equation}
%%
%where $q = (-2)^d \, x_1 \dots x_d$.
%%
%Multiplying \refeq{cbifur} by $q - 1$ yields
%  $1 = q^c = (-2)^n \, x_1 \dots x_n$.
%So the $n$-cycle is at its onset.
%
%
%Further $q$ is a \emph{primitive} $c$th root of unity.
%%
%Suppose the contrary: $q = e^{2k\pi i/c}$ and $(k, c) = g >1$,
%  then by $c_1 \equiv c/g$, $k_1 \equiv k/g$, we have
%\begin{equation}
%  q = e^{2k_1\pi i/ c_1}.
%\label{eq:qd}
%\end{equation}
%Similar to \refeq{cbifur}, we can apply \refeq{stair} with $c \rightarrow g$ and $d \rightarrow d c_1$, and
%\[
%  1 + q_1 + \dots + {q_1}^{g-1} = 0,
%\]
%  where $q_1 = (-2)^{d c_1} \, x_1 \dots x_{dc_1} = q^{c_1}$.
%But by \refeq{qd}, $q^{c_1} = e^{2 k_1\pi i} = 1$,
%and $1 + q_1 + \dots + {q_1}^{g-1} = g > 0$; a contradiction.
%%
%\end{proof}
%
%
%\begin{remark}
%The only real $q$ is $q = -1$ for $c=2$, i.e., a period-doubling.
%%
%On the complex domain, however, we can have a $c$-fold branching with $c>2$,
%  which corresponds to a contact points between ``bulbs''
%  in the inverted Mandelbrot set.
%\end{remark}
%
%By \refthm{cbifur}, $P_n\Rlam$ at the onset point
%includes $P_d(R, e^{2 k \pi i/c} \lambda^{1/c})$ for every
%  possible combination of $k$ and $c$, such that
%  $(k, c) = 1$, $c|n$, and $c > 1$.
%
%Dividing the factors from $P_n\Rlam$ yields \refthm{origfac}.

\subsection{\label{sec:degR}Degrees in $R$}

In the numerical implementation
  (see \refsec{end}),
  it is helpful to know the degrees in $R$ of the polynomials
  to evaluate the determinant by Lagrange interpolation. 
We first show that:
\begin{equation}
\deg_R A_n\Rlam = \textstyle\sum_{d|n} \varphi(n/d) 2^{d-1},
\label{eq:degRA}
\end{equation}
%
The first few values are
1, 3, 6, 12, 20, 42, 70, 144, 270, 540, \dots, from $n =1$.%1034, 2112, 4108, \dots,

From \refeq{xr},
  every term in the expansion of the determinant $|\vct A|$
  can be written as
  $\prod_{p} A_{p,q}$,
  in which, $p$ runs through all cyclic polynomials,
  and $\{q = q(p)\}$ is a permutation of $\{p\}$.
Thus, $\deg_R |\vct A|$ is
  reached by the term that maximizes the product.
Recall that $A_{p,q}$ results from reducing
  $\Lambda(\vx) \, C_p(\vx)$ to $C_q(\vx)$
  by a sequence of substitutions ${x_k}^2 = R-x_{k+1}$.
It follows that to increase $\deg_R$ by 1,
  $\deg_\vx$ is decreased by at least 2.
Accordingly,
  $\deg_R A_{p,q} \le \deg_\vx(\Lambda \, C_p) - \deg_\vx C_q$.
Now running $p$ and $q$ through the $\NB$ cyclic polynomials
  yields
\[
  \deg_R |\vct A | = \deg_R \textstyle\prod_{p} A_{p,q} \le n \, \NB(n),
\]
where we have used $\deg_\vx \Lambda(\vx) = n$.
%
The equality can be achieved, if each $q$ is the ``complement'' of $p$,
  that is, if $\prod_k x_k^{e_k}$ is the generator of $C_p$,
  $C_q$ can be generated from $\prod_k x_k^{1-e_k}$.
For example, for $n=4$,
  $C_1 = x_1 + \cdots + x_4$
  is the complement of
  $C_{123} = x_{234} + \cdots + x_{123}$;
  while $C_{12}$ is its own complement.
Finally, we reach \refeq{degRA} by \refeq{necklace}.

Now, taking the degrees in $R$ of \refeq{primfac} and \refeq{origfac}
  yields
%
\begin{subequations}
\begin{align}
\deg_R P_n\Rlam &\equiv \beta(n) = \sum_{d|n} \mu\Big(\frac{n}{d}\Big) 2^{d-1}, \\
    \deg S_n(R) &= \beta(n)
  - \sum_{d | n, d < n} \beta(d) \, \varphi\Big(\frac{n}{d}\Big).
\end{align}
\label{eq:degR}
\end{subequations}
%
The two formulas are known \cite{mira, stephenson1, blackhurst}.
%
%
The first few
  $\deg_R P_n\Rlam$,
  and $\deg S_n(R)$ are
  1, 1, 3, 6, 15, 27, 63, 120, 252, 495, \dots, %1023, 2010, 4095, \dots,
  1, 0, 1, 3, 11, 20, 57, 108, 240, 472, \dots, % 1013, 1959, 4083, \dots,
  respectively, from $n = 1$.

\begin{table}[h]\footnotesize
  \tbl{
  Smallest positive $r$ at the onset and bifurcation points
  of the $n$-cycles of the logistic map.
  }
{
\begin{tabular}{llll}
%\begin{tabularx}{\linewidth}{
%  >{\hsize=0.5\hsize\centering\arraybackslash}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=0.3\hsize\raggedright\arraybackslash}X
%}
\hline
  $n^\dagger$
& Onset$^\ddagger$
& Bifurcation$^\ddagger$
& \#$^*$ \\
\hline
$1$     & $1.0000000000_1$      &  $3.0000000000_1$       & 1   \\
$2'$    & $3.0000000000_1$      &  $3.4494897428_1$       & 1   \\
$3$     & $3.8284271247_1$      &  $3.8414990075_3$       & 1   \\
$4$     & $3.9601018827_3$      &  $3.9607686524_6$       & 1   \\
$4''$   & $3.4494897428_1$      &  $3.5440903596_6$       & 1   \\
$5$     & $3.7381723753_{11}$   &  $3.7411207566_{15}$    & 3   \\
$6$     & $3.6265531617_{20}$   &  $3.6303887000_{27}$    & 4   \\
$6'$    & $3.8414990075_{3}$    &  $3.8476106612_{27}$    & 1   \\
$7$     & $3.7016407642_{57}$   &  $3.7021549282_{63}$    & 9   \\
$8$     & $3.6621089132_{108}$  &  $3.6624407072_{120}$   & 14  \\
$8'$    & $3.9607686524_{6}$    &  $3.9610986335_{120}$   & 1   \\
$8'''$  & $3.5440903596_{6}$    &  $3.5644072661_{120}$   & 1   \\
$9$     & $3.6871968733_{240}$  &  $3.6872742105_{252}$   & 28  \\
$10$    & $3.6052080669_{472}$  &  $3.6059169323_{495}$   & 48  \\
$10'$   & $3.7411207566_{15}$   &  $3.7425706462_{495}$   & 3   \\
$11$    & $3.6817160194_{1013}$ &  $3.6817266457_{1023}$  & 93  \\
$12$    & $3.5820230011_{1959}$ &  $3.5828117795_{2010}$  & 165 \\
$12'$   & $3.6303887000_{27}$   &  $3.6321857392_{2010}$  & 4   \\
$12''$  & $3.8476106612_{27}$   &  $3.8490363152_{2010}$  & 1   \\
$13$    & $3.6797024578_{4083}$ &  $3.6797038498_{4095}$  & 315 \\
\hline
\multicolumn{4}{p{\linewidth}}{
$^\dagger$
  $\,'$, $\,''$, or $\,'''$ means
    that the cycle is undergoing
    the first, second, or third successive period-doubling, respectively.
} \\
\multicolumn{4}{p{\linewidth}}{
$^\ddagger$
  The subscripts are the degrees of the corresponding minimal polynomial
    of $R = r(r-2)/4$.
} \\
\multicolumn{4}{p{\linewidth}}{
$^*$
  The number of similar cycles.
} \\
\hline
%\end{tabularx}
\end{tabular}
\label{tab:rval}
}
\end{table}

\subsection{\label{sec:end}Implementation}

Let us summarize the algorithm.
%
First, we list \refeqs{xcp}  with $\Lambda(\vx) = \prod_{k=1}^n f'(x_k)$.
This step populates elements of the matrix $\vct T$.
%
The characteristic equation $A_n(\lambda) = \big|\lambda \, \vct I - \vct T\big| = 0$,
with $\lambda$ being $+1$ and $-1$, then gives
the polynomials at onset and bifurcation points, respectively.
%
The contributions from shorter $d$-cycles
  can be removed by
  % computing the characteristic equations for all divisors $d$ and applying
  \refeq{primfac}.
%

When implemented on a computer,
it is often helpful to evaluate the characteristic polynomial
by Lagrange interpolation,
that is, by evaluating $A_n\Rlam$ at a few different $R$,
e.g., $R = 0, \pm1, \pm2,\ldots$, then piecing them together
to a polynomial.
The strategy also allows a trivial parallelization.
%
The algorithm was coded into a Mathematica program,
which was used to compute the polynomials for $n$ up to 13.
%
As shown in \reffig{logtime}, the algorithm
outperformed the resultant \cite{burm} and
  Gr\"obner basis \cite{kk1} methods,
  by at least two orders of magnitudes.
%
Note that the resultant of $f^n(x)-x$ and $\frac{d}{dx}f^n(x)-\lambda$
  is the determinant of the Sylvester matrix, which has
  $2^{n+1}-1$ rows and columns
  [as $\deg_x f^n(x) = 2^n$],
  while the matrix of the characteristic equation \refeq{xr}
  is of only $\NB \approx 2^n/n$ [cf. \refeq{necklace}] rows and columns.
Thus, the current algorithm is expected to be faster.

\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[angle=-90, width=\linewidth]{logtime.pdf}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:logtime}
  Time to compute the $n$-cycle bifurcation polynomials,
    measured from a single core of
    Intel\textsuperscript{\textregistered} 2.5GHz Dual-Core CPU
  (Mathematica version: 8, Magma version: 2.18-4).
  }
\end{figure}
%

%
The exact polynomials of these maps are generally too large to print on paper,
  e.g., the polynomial for the logistic map with $n = 13$
    takes roughly seven megabytes to write down.
%
We therefore save the polynomials and programs on the web site
http://logperiod.appspot.com.
%
The representative $r$ values are listed in \reftab{rval}.

%Compared with other approaches \cite{hitzl, gordon, bailey2, kk1, hao, hao2}
%it appeared to be simpler and more transparent,
%as it only requires the computation of a determinant.

\section{\label{sec:henon}H\'enon map}

A simple two-dimensional extension of the logistic map
  is the H\'enon map \cite{henon}:
\begin{equation}
  x_{k+1} = 1 + y_k - a \, {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\label{eq:henon}
\end{equation}
%
By a change of variables
  $x_k^\mathrm{new} \equiv a \, x_k^\mathrm{old},
   y_k^\mathrm{new} \equiv a \, y_k^\mathrm{old}$,
we have, in terms of the new variables,
\begin{equation}
  x_{k+1} = a + y_k - {x_k}^2, \quad
  y_{k+1} = b \, x_k.
%\tag{$\ref{eq:henon}'$}
\label{eq:henons}
\end{equation}
%
%Since neither $a$ nor $b$ is changed during the transformation,
\refeq{henon} and \refeq{henons} share the
same onset and bifurcation points
in terms of $a$ and $b$.
%
With $b \rightarrow 0$ and $a \rightarrow R$,
we recover the logistic map \refeq{logmaps}.

The second equation of \refeq{henons}
allows us to replace $y_k$ by $b \, x_{k-1}$,
  such that we can work with the cyclic polynomials of $x_k$ only.
The square-free reduction is now
  ${x_k}^2 \rightarrow a + b \, x_{k-1} - x_{k+1}$.

The counterpart of \refeq{der}
  can be found from the Jacobian of $n$th iterate of \refeq{henons}
\[
  \vct J^n
  \equiv
  \left(
    \begin{array}{ccc}
      \partial x_{n+1}/\partial x_1 & \partial x_{n+1}/\partial y_1 \\
      \partial y_{n+1}/\partial x_1 & \partial y_{n+1}/\partial y_1 \\
    \end{array}
  \right),
\]
and the magnitude of both eigenvalues of the matrix must be less than 1.
By the chain rule, we have
  $\vct J^n (x_1) = \vct J^1(x_1) \cdots \vct J^1(x_n)$,
  where
$\vct J^1(x_k)
  =
  \left(
    \begin{array}{cc}
      -2 x_k & 1 \\
      b & 0
    \end{array}
  \right)$.
Thus, the eigenvalues $\lambda$ of $\vct J^n$ can be found from
%
\begin{equation}
\big|\, \lambda \, \vct I - \vct J^n(x_1) \,\big|
  = \lambda^2 - \Theta(\vx) \,\lambda +(-b)^n = 0,
%\tag{$\ref{eq:der}'$}
\label{eq:henonder}
\end{equation}
%
where $\Theta(\vx)$ and $(-b)^n$
  are the trace and determinant of $\vct J^n(x_1)$,
  respectively \cite{hitzl}.
The onset and bifurcation points
  correspond to $\lambda = +1$ and $-1$, respectively

Using the matrix identity $\Tr( \vct A \vct B) = \Tr(\vct B \vct A)$,
  we have
    $\Tr \, \vct J^n(x_1) = \cdots = \Tr \, \vct J^n(x_n)$,
i.e., $\Theta(\vx)$ is a cyclic polynomial of $\vx = \{ x_k \}$.
%e.g., %$\Theta(\vx) = 4 x_1 x_2 + 2 \, b$ for $n = 2$ and
%$\Theta(\vx) = -8 x_1 x_2 x_3 - 2 \, b (x_1+ x_2+ x_3)$ for $n = 3$.
%
Thus, we can expand $\Theta(\vx) \, C_p(\vx)$
  in terms of square-free cyclic polynomials,
  and then replace $\Theta(\vx)$ by $\lambda + (-b)^n/\lambda$
  in the final characteristic equation to obtain the desired polynomials
  (the procedure is similar to that in \refsec{algo} for the logistic map).
%
%
%
The polynomials of $a$ and $b$
  at the onset and bifurcation points for $n$ up to 9
  are listed in the website in \refsec{end}.

\section{\label{sec:cubic}Antisymmetric cubic map}

As an example of a higher-order %one-dimensional
map, we consider the following cubic map \cite{strogatz}
%
\begin{equation}
  x_{k + 1} = f(x_k) = r \, x_k - {x_k}^3.
\label{eq:cubic}
\end{equation}
%
Unlike the logistic map, this map is antisymmetric,
  that is, $f(-x) = -f(x)$.
%
In fact, all antisymmetric cubic maps \cite{brown3, brown4}
  can be mapped to \refeq{cubic} by a change of variables.
We shall show below that this odd parity can be used to reduce the amount of calculation.

First of all, the new replacement rule
\begin{equation}
  {x_k}^3 \rightarrow r \, x_k - x_{k+1}
\label{eq:cubreplace}
\end{equation}
no longer eliminates squares.
%
We must, therefore, extend the basis set of cyclic polynomials
  from the square-free ones to the cube-free ones,
  in listing equations \eqref{eq:xcp},
  e.g.,
  $C_{112} = {x_1}^2 x_2 + \dots + {x_n}^2 x_1$
  is now included in the basis
  (but
  $C_{1112} = {x_1}^3 x_2 + \dots + {x_n}^3 x_1$
  is still not).

The odd parity, however, allows us to limit
  the set of cube-free cyclic polynomials
  to those of even degrees in $\vx$.
%
This is because \refeq{cubreplace} contains only linear and cubic terms,
and a cyclic polynomial with an odd (even) degree in $\vx$
can never be reduced to one with an even (odd) degree.
%
%Thus, we only need polynomials of even degrees.
%
%We therefore have a theorem similar to \refthm{sqrfree}.
%%
%%
%\begin{theorem}
%  For the cubic map \refeq{cubic},
%  any cyclic polynomial $K(\vx)$ of
%  an $n$-cycle orbit
%  $\vct x = \{x_1, \ldots, x_n\}$
%  with an even degree in $\vct x$
%  is a linear combination of
%the even cube-free cyclic polynomials $C_p(\vx)$:
%\[
%  K(\vx) = \sum_{p \in \Pset} f_p(R) C_p(\vx),
%\]
%  where $\Pset = \{0, 11, 12, 13, \ldots, 1122, 1123, \ldots \}$ is
%  the set of indices of all even cube-free cyclic polynomials,
%  and $f_p(R)$ are polynomials of $R$.
%  \label{thm:cubfree}
%\end{theorem}
%
%
%
%
With the above change, the rest derivation is similar to that of the logistic map.
The new $\Lambda(\vx)$ should be
$\prod_{k=1}^n f'(x_k) = \prod_{k=1}^n (r - 3 {x_k}^2)$.
%
The polynomials of $r$
  at the onset and bifurcation points
%for some small $n$ are shown in \reftab{cubpolygen} (general $\lambda$)
%and \reftab{cubpoly} ($\lambda = \pm1$);
%for larger $n$ up to 8,
%we have saved the data
  are listed on the website in \refsec{end}.
%
The representative $r$ values are listed in \reftab{crval}.
%

\begin{table}[h]\footnotesize
%\caption{
\tbl{
  Smallest positive $r$ at the onset and bifurcation points
  of the $n$-cycles of the cubic map.
  }
%\begin{center}
{
\begin{tabular}{l l l l}
%\begin{tabularx}{\textwidth}{
%  >{\hsize=0.5\hsize\centering\arraybackslash}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=0.3\hsize\raggedright\arraybackslash}X
%}
%\begin{tabular*}{\textwidth}{l l l l c | c l l l l}
\hline
  $n^\dagger$
& Onset$^\ddagger$
& Bifurcation$^\ddagger$
& \#$^*$ \\
\hline
$1$     & $1.0000000000_1$      &  $2.0000000000_1$       & 2   \\
$2'$    & $2.0000000000_1$      &  $2.2360679775_2$       & 2   \\
$3$     & $2.4504409645_{8}$    &  $2.4608286739_{12}$    & 4   \\
$4$     & $2.5478350393_{22}$   &  $2.5488312193_{32}$    & 8   \\
$4''$   & $2.2360679775_{2}$    &  $2.2880317545_{32}$    & 2   \\
$5$     & $2.3939250274_{112}$  &  $2.3957922744_{120}$   & 24  \\
$6$     & $2.3334877526_{304}$  &  $2.3355337580_{336}$   & 56  \\
$6'$    & $2.4608286739_{12}$   &  $2.4657090579_{336}$   & 4   \\
$7$     & $2.3729872678_{1080}$ &  $2.3732727868_{1092}$  & 156 \\
$8$     & $2.3525990555_{3108}$ &  $2.3527637793_{3200}$  & 400 \\
$8'$    & $2.5488312193_{32}$   &  $2.5493247379_{3200}$  & 8   \\
$8'''$  & $2.2880317545_{32}$   &  $2.2992279397_{3200}$  & 2   \\
\hline
\multicolumn{4}{p{\linewidth}}{
$^\dagger$
  $\,'$, $\,''$, or $\,'''$ means that the cycle is undergoing
    the first, second, or third successive period-doubling, respectively.
} \\
\multicolumn{4}{p{\linewidth}}{
$^\ddagger$
  The subscripts are the degrees of the corresponding minimal polynomial.
} \\
\multicolumn{4}{p{\linewidth}}{
$^*$
  The number of similar cycles
    (for $n > 1$, only half of them have positive $r$).
} \\
\hline
\end{tabular}
%\end{tabularx}
%\end{tabular*}
%\end{center}
\label{tab:crval}
}
\end{table}

%Similar to the logistic-map case,
Even cube-free cyclic polynomials
can be mapped to \emph{even ternary} necklaces,
in which each bead can assume
0, 1, or 2, with the total being even.
%
The number of ternary necklaces is
$N_3(n) = \frac{1}{n} \sum_{d|n} \varphi(n/d) 3^d$,
%\begin{theorem}
while the even ones sum to
%or the cube-free cyclic polynomials for the cubic map of even degrees in $\vx$
%is given by
\begin{equation}
  N_e(n) = \frac{1}{n} \sum_{c d = n} \varphi(c)
    \left[
      3^d - \odd(c) \frac{3^d-1}{2}
    \right],
\label{eq:cubnecklace}
\end{equation}
where
$\odd(c)$ %$\equiv \big[1 - (-1)^c\big]/2$
is 1 if $c$ is odd or 0 if even.
%\label{thm:cubnecklace}
%\end{theorem}
%
Since $N_e \approx N_3/2$, using the parity saves about half of the equations
  in computation.
%
$N_e(n) = $ 2, 4, 6, 14, 26, 68, 158, 424, \dots, from $n = 1$.
%

The characteristic polynomial $A_n\rlam$
  from the determinant equation
  has a degree $N_e(n)$ in $\lambda$.
Again, it encompasses the factors
  from the shorter $d$-cycles ($d|n$),
  and the $n$-cycle polynomial can be obtained
  from \refeq{primfac} by proper substitutions, e.g.,
  $R\rightarrow r$.
%The degree of the polynomial,
The degree in $\lambda$ of $P_n\rlam$
  is given by, after some algebra,
%
%
\begin{equation}
  L_e(n) = \frac{1}{n} \sum_{cd = n} \mu(c)
    \left[
      1 + \odd(c)\frac{3^d-1}{2}
    \right],
\label{eq:cublyndon}
\end{equation}
%
%
For the first few values,
  $L_e(n)$ = 2, 2, 4, 10, 24, 60, 156, 410, \ldots, starting from $n = 1$.
%
This is also the number of the $n$-cycles \cite{hao, hao2}.
Note that, for $n>1$, $x_k$ in half of the cycles
  with negative $r$ are imaginary.
However, in a transformed map
$z_{k+1} = r \, z_k \, (1 - {z_k}^2)$,
which differs from \refeq{cubic} by $x_k = \sqrt{r} z_k$,
$z_k$ in the negative-$r$ cycles are real.
%
%
Following similar steps in \refsec{degR},
we find that
  $\deg_r A_n(r, \lambda) = n N_e(n)$
  and $\deg_r P_n(r, \lambda) = n L_e(n)$.

%\subsection{\label{sec:oddcycle}Odd-cycles}

%
%
\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[angle=-90, width=\linewidth]{oddcycle.pdf}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:oddcycle}
  Odd-cycles of the cubic map.}
\end{figure}
%
%

The symmetry $f(-x) = -f(x)$ permits
  a particular \emph{odd-cycle},
  see \reffig{oddcycle} for examples.
If $x_{n+1} = -x_1$ has a solution,
  then $x_1, \ldots, x_{n}, -x_1, \ldots, -x_{n}$
  is a $2n$-cycle, for $x_{2n+1} = - x_{n+1} = x_1$.
%
It follows that a cyclic polynomial
  formed by the $2n$ points
  with an odd degree in $\vxn{2n}$
  is also always zero,
  e.g., $C_1 = x_1 + x_2 + \dots + x_n - x_1 - x_2 - \dots - x_n = 0$.
Accordingly, the characteristic equation \refeq{xr}
  would omit the $n$-odd-cycles from the solution.
The existence of odd-cycles is why we use even, but not odd,
  cyclic polynomials in the solution.

Since the $n$-odd-cycles satisfy a polynomial of lower degree,
  the $n$-cycle polynomial $P_{2n}(\lambda)$
  (omitting the argument $r$)
  can always be factorized.
%
Suppose $\Lambda^o(\vx) \equiv \prod_{k = 1}^{n} f'(x_k)$ in the odd-cycle
 satisfies $P_{n}^o(\lambda^o) = 0$
\big[where $\lambda^o$ is the value of $\Lambda^o(\vx)$,
and $\lambda^o = \pm\sqrt \lambda$\,\big],
then $P^o_{n}(\sqrt \lambda) P^o_{n/2}(-\sqrt \lambda)$
is a factor of $P_{n}(\lambda)$.
%
For example, by solving the $n = 2$ odd-cycle,
  we have $- x_1 = r x_1 - {x_1}^3$, or ${x_1}^2 = r + 1$.
Since $\lambda^o = r - 3 {x_1}^2$,
$P^o_1(\lambda^o) = \lambda^o + 2 r + 3$.
%
The 2-cycle polynomial
  $P_2(\lambda) = -\big[(2r+3)^2 - \lambda\big]\,(\lambda + 2r^2-9)$,
  indeed contains
  $P^o_1(\sqrt \lambda) \, P^o_1(r, -\sqrt \lambda)
   = (\sqrt \lambda + 2 r + 3)(-\sqrt \lambda + 2 r + 3)$
  as a factor.

Finally, we mention that the $n$-cycle polynomial $P_n\rlam$
  is an even polynomial of $r$ if $n$ is a multiple of 4.
%This follows from that the transformation
%  $x_k \equiv (-)^k i y_k, r \equiv -r'$
%  yields the identical map: $y_{k+1} = r' y_k - {y_k}^3$.
Practically, we can cut the computating time by half
  when calculating the determinant by Lagrange interpolation
  in this case.

\nonumsection{Acknowledgments} \noindent
It is a pleasure to thank T. Gilbert, Y. Mei and the referees
  for helpful communications and suggestions.
Computing time on the Shared University Grid at Rice,
  funded by NSF under Grant EIA-0216467, is gratefully acknowledged.

%\appendix{}

% ws-ijbc dislikes the following
%\appendix

%\appendixpage

%\appendix{\label{apd:per4}Derivation of the 4-cycles}
%
%
%
%The polynomials for the 4-cycles has a compact derivation.
%We first list the explicit equations:
%%
%%
%%
%\begin{subequations}
%\label{eq:x4}
%\begin{align}
%  x_2 &= R - {x_1}^2, \\
%  x_3 &= R - {x_2}^2, \\
%  x_4 &= R - {x_3}^2, \\
%  x_1 &= R - {x_4}^2,
%\end{align}
%\end{subequations}
%%
%$\big[\mathrm{\refeqsub{x4}{a}} - \mathrm{\refeqsub{x4}{c}}\big]
%\times
%\big[\mathrm{\refeqsub{x4}{b}} - \mathrm{\refeqsub{x4}{d}}\big]$
%yields $1 + (x_1 + x_3) (x_2 + x_4) = 0$,
%since $x_1 \ne x_3, x_2 \ne x_4$.
%%
%Hence, with
%$y_1 \equiv x_1 + x_3$, $y_2 \equiv x_2 + x_4$,
%$z \equiv y_1 + y_2$,
%we have
%%
%%
%%
%\begin{subequations}
%\begin{align}
%y_1 y_2           &= -1, \\
%{y_1}^2 + {y_2}^2 &= z^2 - 2 \, y_1 y_2 = z^2 + 2, \\
%{y_1}^3 + {y_2}^3 &= z^3 - 3 \, y_1 y_2 \, z = z^3 + 3 z.
%\end{align}
%\label{eq:ypow4}
%\end{subequations}
%%
%%
%Multiplying \refeqsub{x4}{a} by $x_1$ or $x_3$,
%then summing over cyclic versions yields
%\begin{subequations}
%\begin{align}
%y_1 y_2 &= R z - \big[({x_1}^3 + {x_3}^3) + ({x_2}^3 + {x_4}^3)\big],\\
%y_1 y_2 &= R z - \big[x_1 x_3 \, y_1 + x_2 x_4 \, y_2\big].
%\end{align}
%\label{eq:p4q}
%\end{subequations}
%%
%From
%$\mathrm{\refeqsub{p4q}{a}} + 3 \times \mathrm{\refeqsub{p4q}{b}}$,
%we have
%$4 \, y_1 y_2 = 4 R z - ({y_1}^3 + {y_2}^3)$,
%%
%and by \refeqs{ypow4},
%\begin{equation}
%  z^3 - (4 R - 3) z - 4 = 0.
%  \label{eq:xr4s}
%\end{equation}
%%
%%
%Since
%$2 x_1 x_3 = {y_1}^2 - ({x_1}^2 + {x_3}^2)
%= {y_1}^2 - 2 R + y_2$,
%and $2 x_2 x_4 = {y_2}^2 - 2 R + y_1$,
%%
%%Using \refeqs{ypow4} to simplify the right hand side, we find
%\begin{equation}
%  X \equiv x_1 x_2 x_3 x_4 = \tfrac{1}{2} R z(1 - z) + (R^2 - R + 1),
%\label{eq:der4}
%\end{equation}
%where we have used \refeqs{ypow4} and \req{xr4s} to simplify the result.
%%$X=1/16$ and $-1/16$ at the onset and bifurcation points, respectively.
%%
%Dividing the polynomial in \refeq{xr4s} by that in \refeq{der4}
%yields $z = (R^2-3R-X+1)/(R^2-R+X-1)$,
%and plugging it back to \refeq{der4} gives
%   $R^6 -3 R^5
%  + (3 + X) (R^4 - R^3)
%  + (1 - X) (2 + X) R^2
%  + (1 - X)^3 = 0$,
%which is the same as the sextic factor of \refeq{xr4}
%  with $X = \lambda/16$.
%
%

%\section{\label{apd:cublyndon}Proof of \refthm{cublyndon}}

\begin{thebibliography}{50}

\bibitem[May(1976)]{may}
  May, R. H. [1976]
  ``Simple mathematical models with very complicated dynamics,''
  {\it Nature}
  \textbf{261},
  459--467.

\bibitem[Strogatz(1994)]{strogatz}
  Strogatz, S. H. [1994]
  {\it Nonlinear Dynamics and Chaos}
  (Addison-Wesley. Reading, MA).

\bibitem[Brown(1981)]{brown1}
  Brown, A. [1981]
  ``Equations for periodic solutions of a logistic difference equation,''
  {\it J. Austral. Math. Soc. Ser. B}
  \textbf{23},
  78--94
  (1981).

\bibitem[Brown(1982)]{brown2}
  Brown, A. [1982]
  ``Solutions of period seven for a logistic difference equation,''
  {\it Bull. Austral. Math. Soc.}
  \textbf{26},
  263--284.

\bibitem[Stephenson(1991)]{stephenson1}
  Stephenson, J. [1991]
  ``Formulae for cycles in the Mandelbrot set,''
  {\it Physica A}
  \textbf{177},
  416--420.

\bibitem[Stephenson(1992a)]{stephenson2}
  Stephenson, J. [1992a]
  ``Formulae for cycles in the Mandelbrot set II,''
  {\it Physica A}
  \textbf{190},
  104--116.

\bibitem[Stephenson(1992b)]{stephenson3}
  Stephenson, J. [1992b]
  ``Formulae for cycles in the Mandelbrot set III,''
  {\it Physica A}
  \textbf{190},
  117--129.

\bibitem[Saha \& Strogatz(1995)]{saha}
  Saha, P. \& Strogatz, S. H. [1995]
  ``The birth of period three,''
  {\it Mathematics Magazine}
  \textbf{68},
  42--47.

\bibitem[Bechhoefer(1996)]{bechhoefer}
  Bechhoefer, J. [1996]
  ``The birth of period 3, revisited,''
  {\it Mathematics Magazine}
  \textbf{69},
  115--118.

\bibitem[Gordon(1996)]{gordon}
  Gordon, W. B. [1996]
  ``Period three trajectories of the logistic map,''
  {\it Mathematics Magazine}
  \textbf{69},
  118--120.

\bibitem[Burm \& Fishback(2001)]{burm}
  Burm, J. \& Fishback, P. [2001]
  ``Period-3 orbits via Sylvester's theorem and resultants,''
  {\it Mathematics Magazine}
  \textbf{74},
  47--51.

\bibitem[Zhang(2010)]{zhang}
  Zhang, C. [2010]
  ``Period three begins,''
  {\it Mathematics Magazine}
  \textbf{83},
  295--297.

\bibitem[Bailey \& Broadhurst(2000)]{bailey1}
  Bailey, D. H. \& Broadhurst, D. J. [2000]
  ``Parallel integer relation detection: techniques and applications,''
  {\it Mathematics of Computation}
  \textbf{70},
  1719--1736.

\bibitem[Bailey {\it et al.}(2006)]{bailey2}
  Bailey, D. H., Borwein, J. M., Kapour, V. \& Weisstein, E. W. [2006]
  ``Ten problems in experimental mathematics,''
  {\it The American Mathematical Monthly}
  \textbf{113},
  481--509.

\bibitem[Kotsireas \& Karamanos(2004)]{kk1}
  Kotsireas, I. S. \& Karamanos, K. [2004]
  ``Exact computation of the bifurcation point $B_4$ of the logistic map
  and  the Bailey-Broadhurst conjectures,''
  {\it International Journal of Bifurcation and Chaos}
  \textbf{14},
  2417--2423.

\bibitem[Lewis(2008)]{lewis}
  Lewis, R. H. [2008]
  ``Heuristics to accelerate the Dixon resultant,''
  {\it Mathematics and Computers in Simulation}
  \textbf{77},
  400--407.

\bibitem[Mandelbrot(1980)]{mandelbrot}
  Mandelbrot, B. B. [1980]
  ``Fractal aspects of the iteration of $z \rightarrow \lambda z(1-z)$
      for complex $\lambda$ and $z$,''
  {\it Ann. NY Acad. Sci.}
  \textbf{357},
  249--259.

\bibitem[Hardy \& Wright(2008)]{hardy}
  Hardy, G. H. \& Wright, E. M. [2008]
  {\it An Introduction to the Theory of Numbers}
  (Oxford University Press, USA).

\bibitem[Hao(1989)]{hao}
  Hao, B.-L. [1989]
  {\it Elementary Symbolic Dynamics and Chaos in  Dissipative Systems}
  (World Scientific, Singapore).

\bibitem[Hao(2000)]{hao2}
  Hao, B.-L. [2000]
  ``Number of periodic orbits in continuous maps of
  the interval complete solution of the counting problem,''
  {\it Annals of Combinatorics}
  \textbf{4},
  339--346.

\bibitem[Lutzky(1988)]{lutzky}
  Lutzky, M. [1988]
  ``Counting stable cycles in unimodal iterations,''
  {\it Physics Letters A}
  \textbf{131},
  248--250.

\bibitem[Mira(1987)]{mira}
  Mira, C. [1987]
  {\it Chaotic Dynamics}
  (World Scientific, Singapore).

\bibitem[Blackhurst(2011)]{blackhurst}
  Blackhurst, J. [2011]
  ``Polynomials of the bifurcation points of the logistic map,''
  {\it International Journal of Bifurcation and Chaos}
  \textbf{21},
  1869--1877.

\bibitem[H\'enon(1976)]{henon}
  H\'enon, M. [1976]
  ``A two-dimensional mapping with a strange attractor,''
  {\it Communications in Mathematical Physics}
  \textbf{50},
  69--77.

\bibitem[Hitzl(1985)]{hitzl}
  Hitzl, D. L. \& Zele, F. [1985]
  ``An exploration of the H\'enon quadratic map,''
  {\it Physica D}
  \textbf{14},
  305--326.

\bibitem[Brown(1984a)]{brown3}
  Brown, A. [1984a]
  ``Solutions of period three for a non-linear difference equation,''
  {\it J. Austral. Math. Soc. Ser. B}
  \textbf{25},
  451--462.

\bibitem[Brown(1984b)]{brown4}
  Brown, A. [1984b]
  ``Solutions of period four for a non-linear difference equation,''
  {\it J. Austral. Math. Soc. Ser. B}
  \textbf{26},
  146--164.

\end{thebibliography}

\end{document}